--- 
title: "Reading Notes"
author: "Hui Lin"
date: "`r Sys.Date()`"
knit: "bookdown::render_book"
documentclass: krantz
bibliography: [bibliography.bib]
biblio-style: apalike
link-citations: yes
colorlinks: yes
lot: yes
lof: yes
fontsize: 12pt
always_allow_html: yes
monofont: "Source Code Pro"
monofontoptions: "Scale=0.7"
site: bookdown::bookdown_site
description: "Hui's Reading Notes"
url: 'http\://hui1987.com'
github-repo: happyrabbit/IntroDataScience
---

```{r include=FALSE, echo = FALSE}
# Sys.setenv(TZ="UTC")
options(
  htmltools.dir.version = FALSE, formatR.indent = 2,
  width = 55, digits = 4, warnPartialMatchAttr = FALSE, warnPartialMatchDollar = FALSE
)

# ids_url = function(file) {
#  paste0('https://github.com/',
#         'happyrabbit/DataScientistR/',
#         'raw/master/', file)
# }

#bookdown::render_book("index.Rmd", "bookdown::gitbook")
#bookdown::render_book("index.Rmd", "bookdown::pdf_book")
```




<!--chapter:end:index.rmd-->

\mainmatter


# Introduction

Learning is happening increasingly outside of formal educational settings and in unsupervised environments. Technological advantages provide new tools and opportunities for life-time learners. This reading notes is my first step of managing my learning (aka: Personal Knowledge Management). This step is called Personal Information Management(PIM). The goals are:

1. Organize information
1. Initially process the information and understand the information structure
1. Chunking: cluster similar information, filter out some detail, degrade the complexity

The reading notes are arranged according to book category in my reading list since 2009. For some reason I lost all reading information before 2009. There are tons of paper notes need to be digitized and I will keep working on it in my spare time. My personal comments will start with “Hui:” in blue. Sometimes I summarize the contents with my own words. It is great if you find any of those notes useful for you as well. I thank all my friends who recommended great books to me. I wrote this notes with [R Markdown](https://rmarkdown.rstudio.com)  and the R package [bookdown](https://bookdown.org/yihui/bookdown/).

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(DT)
library(ggplot2)
dat <- read.csv("1_KnowledgeStructure/RawData/ReadingListHui.csv")
dat %>%
  select(Name, ProDetail, Category, Topic) %>%
  datatable(filter = "top", options = list(pageLength = 50))
```

- **Name**: Book name
- **Category**: Book category
- **ProDetail**: My reading status
    - `HaveNotRead`: have not started reading
    - `RoughReading`: just got the structure and key ideas of the book without getting to every glorious detail
    - `PartialReading`: read carefully for some parts of the book but not the whole
    - `ThroughReading`: read word by word for the whole book
    - `IntensiveReading`: read repeatedly and try to understand it very well
- **Topic**: the subject of the book. I use the information to classify books to do “syntoptical reading” which is to read two or more books on the same subjects. You may find lots of missing values here as I just began my journey on this. Also, it is not easy to identify books with “same subject”. I will only assign subject to non-fictions.
- **Priority**: priority of writing book summary

The categories are:

```{r}
dat %>%
  group_by(Category) %>%
  summarise( count = n()) %>%
  arrange(desc(count)) %>%
  datatable()
```

## Rules of Generating book ID

It is tentative. I don’t think I will label my books. It was originally motivated by being able to find a book more easily. But it seems that I remember where to get a book when I need it.

1. **Category code** + **Number**
2. **Number** was origionally assigned as the rank of the book in its category devided by 100 (i.e rank/100)
3. For new book
    1. search for the maximum number in that category (`mn`)
    2. use `mn+0.01` as number for the new book

## Basic Statistical Analysis

Here is a plot for counts of books in each category……

```{r}
pdat <- dat %>%
  group_by(Category) %>%
  summarise(Count = n())
ggplot(pdat, aes(x = reorder(Category, Count), y = Count, fill = Category)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  theme(legend.position="none", axis.title.y=element_blank()) 
```

Here is a plot showing the percentage of books I have started (or finished) across different categories.

```{r}
ggplot(dat,aes(Category, fill = ProDetail)) +
  geom_bar(position = "fill") +
  coord_flip() +
  theme( axis.title.y=element_blank(),
         axis.title.x=element_blank()) 
```


<!--chapter:end:01-Introduction.Rmd-->

# Cognitive Science

## Thinking Fast and Slow

一、大脑的运作机制

1、大脑的运作是双核系统，有快与慢两种系统模式。

系统1：快的思考方式，即直觉系统，就像自动驾驶系统，指人类无意识的快速思考模式，依赖情感、经验和记忆来做决策判断，不怎么费脑力。

系统2：慢的思考方式，就像手动驾驶，指有意识进行的一种更慢、更严谨，需要投入更多脑力的思考模式，不容易出错但它很懒惰，经常走捷径，直接采纳系统1的直觉型判断结果。

2、误区：主导决策和判断的主角是系统2？

其实真正的主角是系统1，系统2只是配角。

因为大多数情况下都是系统1在处理，而系统1用直觉进行思考判断的速度非常快，快到往往在人们还没有意识到就将问题处理完了。而如果让系统2来处理，凡事都通过深思熟虑来做决策，实在是太耗费脑力和精力。

3、系统1和系统2的运作机制

这两个系统之间的互动，构成了大脑认知和思考的运作机制：

大多数情况下是系统1用直觉来快速处理，只有在系统1遇到解决不了的问题时，才求助于系统2，系统2才会调动注意力、分析力来思考和解决问题。

优势：这样既能节约脑力，又能解决绝大多数问题。

弊端：我们容易被系统1也就是直觉所欺骗和糊弄。

原因是这种互动的运作机制存在漏洞，即有时候系统1面对不擅长解决而需要系统2介入的问题，却非要用直觉来做判断和处理，而系统2容易怠倦就盲目相信系统1能处理，最终导致决策失误。

二、欺骗我们的三大直觉漏洞

直觉系统会对我们的观点和行为有无所不在的影响，却存在不可避免的错误和偏颇，导致我们容易被直觉糊弄和欺骗，从而过分自信和决策失误。

直觉漏洞一：简单联想
指做决定并不是深思熟虑根据分析和判断来进行，而是直觉系统启动联想效应，以联想到的相关事情来作为判断的重要依据。

启动效应

也可以说联想效应，指大脑一旦接受一个新的概念或者信息的时候，立刻启动联想这个机制，带动与此相关的其他记忆和感情。

案例一：投票地点对投票结果的影响

案例二：图像对自助付款的监督效果对比

案列三：阅读包含老人意象的词组后影响行动力变慢

案例四：慢速走路后更容易发现类似老人这样的词组

很多人利用简单联想来影响我们的行为和决策

广告商深谙简单联想的巨大威力，比如可口可乐。

我们自己因简单联想的影响而决策失误

很可能在生活中因为前一次成功的经验就会简单联想到之后也会成功。

比如：赌钱的人，赢钱了就觉得接下来还会赢。

直觉漏洞二：易得性判断

指只利用大脑能够想得起来的信息进行判断。直觉系统的这个漏洞，会给决策带来很多问题。比如：100个因素需要分析，大脑只能想到5个因素，直觉就会只用这5个因素来拼凑一个合理的结论。

故事比统计概率更能说服我们。

直觉系统，宁愿相信貌似合理的结论而不是一个可能性更高的结论。这和逻辑、统计概率相矛盾

案例一：心理学经典案列“琳达实验”

易得性原理: 指直觉系统通过易得性的数据信息进行貌似合理的分析判断，而不容易得到的信息，都不会主动去思考判断，因为快思考的模式，而容易忽略重要的关键因素。

比如：因为学生的纹身而判断他成绩不好。因为领导演讲很棒，就觉得他是一个合格的好领导。

直觉漏洞三：因果联系

指直觉系统习惯于用因果关系解释自己观察到的现象，否认随机性。

对随机性的误解和错觉。未经过统计学训练的人，很难相信随机性，更倾向于用虚假的因果关系去解释他们认为不随机的事件。

案例一：篮球运动中的投篮顺手观点，其实只是随机现象带来的“运气”

在随机事件上寻找因果联系会导致极大的损失

案例二：小样本带来的统计大损失——比尔盖茨基金会赞助17亿美元的无用研究

三、如何避免被直觉的漏洞所欺骗？

解决方案很简单，即在关键时候让系统二介入来处理。
难点在于：质疑直觉系统很难，会让我们有压力，不愉快，想要逃避。这也是为什么直觉系统这些漏洞发生后，我们根本意识不到的原因。

两个实用方法

方法一：引入外部意见

指旁观者清，应当多听从外部的信息来做思考判断。

案例一：作者花费8年事件编写出版教材的失败经历

方法二：核查清单

相当于给大脑找一个外部专家，借助清单这样的外部工具，将直觉带来的错觉写进核查清单来提醒自己，在关键时刻抵制直觉系统犯错误。

案例一：查理芒格用误判清单来提醒自己在重大投资和决策的时候不要犯错。

<!--chapter:end:02-PHY.Rmd-->

# Self-development
## Emotion Management

### Emotional First Aid (情绪急救)

> Dr. Guy Winch is a licensed psychologist, author, and in-demand keynote speaker whose books have been translated into 24 languages. His first TED Talk, Why We All Need to Practice Emotional First Aid has been viewed over 5 million times and is rated as the #5 most inspiring TED Talk of all time on ted.com He also writes the popular Squeaky Wheel Blog on Psychology Today.com (https://www.psychologytoday.com/blog/the-squeaky-wheel) and he has dabbled in stnad-up comedy. His website is www.guywinch.com. You can view his TED Talks here: https://www.ted.com/speakers/guy_winch

Healing Rejection, Guilt, Failure, and Other Everyday Hurts

### How To Keep People From Pushing Your Buttons

> About the author: Albert Ellis(1913-2007) held M.A and Ph.D. degrees in Clinical Psychology from Columbia University. He was the founder of Rational Emotive Behavior Therapy (REBT), the pioneering form of the modern Cognitive Behavior therapies. He was the president of the Albert Ellis Institute in New York, where he practiced individual and group psychotherapy, supervised and trained psychotherapists, and presented many talks and workshops at the Institute and throughout the world. He published over seven hundred articles and more than sixty books on psychotherapy, marital and family therapy, and sex therapy.

4 fatal "Overly":

1. Anxious
1. Angry
1. Depressed
1. Guilty

3 nutty beliefs (Realistic Preferences):

(1) Awfulizing
(2) Shoulding
(3) Rationalization

The three nutty beliefs will further lead to 10 nutty beliefs:

(1) Worry too much about what other people think of you
    - Please others
    - Overly defensive

(2) Excessively fear of failure
    - Lack of risk-taking
    - Stagnation
    - Mediocrity

> If you are hitting the bullseye every time you shoot, you are standing too close!

(3) Low frustration tolerence
    - impulsive reactions, premature quitting, nagativism, avoidance of responsibility, lack of perseverance
    
(4) Blame others

(5) Worry obsessively about things to happen

(6) Perfectionism

(7) Avoid responsibility

(8) Detachment

(9) Blame past experience

(10) Bad things/people shouldn't exist

<!--

Ellis列出了四种不好的过激情绪：
过分的烦躁（紧张、沮丧、恼火、担惊受 怕）；比如你为准备一个工作面试而紧张，为孩子最近的表现而烦躁。
过分生气（戒备、被激怒，气得发疯，挫 败）：当你的配偶批评你不干活，当工作上的同事不配合或做不来时。
过分抑郁（无精打采，一蹶不振）：失去所爰的人，或失去工作。
过分内疚（过分承担责任，过分悔恨，过分自责）：比如你因为离婚而对孩子内疚，处处宠溺。
如何判断过分？其实人们在85%的时间里，是能识别出自己什么时候反应过激。而你要做的就是让尽可能多的过激反应不要发生。

你的坏情绪是怎么来的？人不是天生就有坏情绪的，Ellis给出了诱因ABC公式。
A代表曰常遇见的人或事情。除了那些重大的事件外，A就是那些烦人的小事，比如难缠的 上司、同事，办公室的勾心斗角，沟通不够； 生活中与配偶或情人的冲突，家务劳动，财务 问题。

C代表在A的情形下你的感觉和行为。如果你有一个重要的会议，但路上大堵车，这时你就会从平时的绅士变成路怒一族，不断并线加塞，嘴里骂骂咧咧。那么真的是堵车导致了你这样的路怒行为了吗？不是的，在A和C之间，有一个B

B就是我们对具体发生的人或 事的思考、判断。

这样你就会明白，不是A这件事一定会产生C的结果，A本身不会导致C，而是B导致了C，A只是诱因。

三类病态的思维模式导致情绪和行为反应过激
恐怖化的思维方式。把什么都看成了灾难，非常害怕，思维模式就是“万一……怎么 办”？比如你去面试一份工作，你可能会想 到：万一我回答不上问题怎么办？万一他们不 喜欢我怎么办？万一我被人难堪了怎么办？越想越多，就越恐慌，你高度紧张，未战而败。

应该化的思维方式。“我必须……’”我一定……“，这种”必须”“应该”的思维可能源自孩提时代，父母说：你应该对弟弟好一点；老师说：你应该五讲四美。一旦有了 “应该”的思维，就会使你对自己要求过于严苛，把自己弄得很惨。当你达到了我应该“的标准时，你就开始了对别人的”你应该”。

合理化的思维方式。这种思维就是认为什么都很合理，不可思议地将不道德或不得体的 行为合理化，骗自己接受这种行为，简单地说 就是逆来顺受。典型的表现就是：谁会关心？ 那又怎样？这种思维方式会把很多不合理的事情也认为是合理的，合理化思维是一种软弱的应对方式。

你如果老是用恐怖化、应该化、合理化去思考 问题的话，就会陷入以下10种非理性人生信条：

信条1:太在乎别人怎么看待你。因为太在乎 会导致你对拒绝的强烈的恐惧感，要么你会四处讨好别人，避免冲突；要么你就把自己装扮 成一只刺猬，逮谁扎谁。其实你更好的选择可以是这样的：“我希望你们喜欢我，尊重我， 你们要是做不到，我也能忍受”。（恐怖化）

信条2:无法忍受在重要任务上失败。如果你让自己相信在任何重大事件上都输不起，你就 不敢冒险，会墨守成规。还有就是无法忍受别人的批评。（恐怖化）

信条3:人和事都应该朝着我要的方向发展。这是由于低耐挫性和对不公平的敏感而造成的 冲动反应，会导致半途而废，逃避责任。（应该化）

信条4:某件事出错了，肯定是有人出了问题。凡事绝对化，认为事情做得那么糟，肯定 是有人不尽力，这些人善于把脏水往别人身上泼。（应该化）

信条5:对即将发生的事情总是抱着深深的忧虑。（恐怖化）

信条6:每个问题都有完美的解决方案，我必须立即找到这些方法。有时候完美主义造就了拖延症。（应该化）

信条7:在很多困境和责任面前可以让自己置身事外，凡事都能找到一个合理化的理由。 (合理化）

信条8:如果事事不投入，保持若即脱离的关注，会永远开心。很多人貌似参与了很多事， 但其实他们只是坐在那儿，被动地观察。（合理化）

信条9:因为过去或者小时候发生的一些不好的事情造成了我现在这个样子，我努力也改变不了。（合理化）

信条10:这些坏人坏事就不应该存在，真不知道拿他们怎么办。他人他事会逼得你反应过激。（应该化）

怎样应对病态的思维方式？
把事情恐怖化、应该化和合理化是三大逼迫我们就范的方式，要我们不去这么想的方法就是 第四种思考模式：你要看看在这三种思维模式 之外有什么是更好的选择，比如“我想 要……，”我更喜欢……“，”如果……就更好了“。它不是必须时刻要积极，并不表明你能够或一定成功，而是说要去试一试，即使可能会失败、被拒绝。

比如你是个演讲新手，但最近你有一场重要的 演讲，你很可能会把前面三类病态思维都用一 下：“万一我搞砸了，没人听，大家聊天会怎 么办？（恐怖化）我应该能够做好公共演讲， 我用不着害怕。（应该化）其实，这有多大的事，就是搞砸了也没人在乎。（合理化）”但如果你能更好的选择思维方式：嗯，我想要做好演讲，但也不是非做好不可，现在还有一点时间，我可以在脑子里再过一次提纲；其实我更喜欢演讲的过程，我想我肯定能从中学到什么。这样你就不会背负心理包袱。

如何让“更好的选择”成为习惯？
改变想法需要觉悟，还有需要用大量的练习建 立起第四种思维模式的习惯‘‘更好的选择“。 有四个步骤：

反思自己的C (感觉和行为）：我的感觉和行为是不是合适？
认真审视自己的B是怎么把自己弄成了C的样子（过分的焦虑、愤怒、抑郁、内疚）
如何反击和对抗自己的非理性思考方式
用什么样的更好选择来替代非理性思维
用“我想要……”，“我喜欢……”，“如果……就更好了”
--!

### Others

- From TED talk: [How to fix a broken heart](https://www.ted.com/talks/guy_winch_how_to_fix_a_broken_heart?utm_source=newsletter_weekly_2018-02-10&utm_campaign=newsletter_weekly&utm_medium=email&utm_content=talk_of_the_week_image)

Heartbreak is far more insidious than we realize. There is a reason we keep going down one rabbit hole after another, even when we know it's going to make us feel worse. Brain studies have shown that the withdrawal of romantic love activates the same mechanisms in our brain that get activated when addicts are withdrawing from substances like cocaine or opioids.

Getting over heartbreak is not a journey. It is a fight, and your reason is your strongest weapon. There is no breakup explanation that is going to feel satisfying. No rationale can take away the pain you feel. So don't search for one. Don't wait for one, just accept the one you were offered or make up one yourself and then put the question to rest, because you need that closure to resist the addition. And you need something else as well: you have to be willing to let go, to accept that it's over. Otherwise, your mind will feed on your hope and set you back. Hope can be incredibly destructive when your heart is broken.

<!--chapter:end:13-SDP.Rmd-->

