---
title: "R HTMLWidgets"
output: flexdashboard::flex_dashboard
---

Page 1
===================================== 
    
### `Dygraphs`: provides rich facilities for charting time-series data in R and includes support for many interactive features.
    
```{r}
library(dygraphs)
dygraph(sunspots) %>% dyRangeSelector()
```
    
### `Plotly`: translate your ggplot2 graphics into an interactive web-based version.

```{r}
library(plotly)
p <- ggplot(data = diamonds, aes(x = cut, fill = clarity)) +
            geom_bar(position = "dodge")
ggplotly(p)
```

Page 2
=====================================     

### `d3heatmap`: creates interactive D3 heatmaps including support for row/column highlighting and zooming.

```{r}
library(d3heatmap)
d3heatmap(mtcars, scale="column", colors="Blues")
```
    
### `MetricsGraphics`: enables easy creation of D3 scatterplots, line charts, and histograms.

```{r}
library(metricsgraphics)
mjs_plot(mtcars, x=wt, y=mpg) %>%
  mjs_point(color_accessor=carb, size_accessor=carb) %>%
  mjs_labs(x="Weight of Car", y="Miles per Gallon")