---
title: "Tidy and Reshape Data"
author: Hui Lin
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

There are two commonly used packages for this kind of manipulations: `tidyr` and `reshape2`. 

## `reshape2` package

It is a reboot of previous package reshape.

Take a baby subset of our exemplary clothes consumers data to illustrate:


```{r, message=FALSE}
library(readr)
library(reshape2)
library(tidyr)
sim.dat <- read_csv("https://raw.githubusercontent.com/happyrabbit/DataScientistR/master/Data/SegData.csv ")
(sdat<-sim.dat[1:5,1:6])
```

## Linear Regression

You can run a regression to study the effect of purchasing channel:

```{r}
# Here we use all observations from sim.dat
mdat<-melt(sim.dat[,1:6], measure.vars=c("store_exp","online_exp"),
            variable.name = "Channel",
              value.name = "Expense")
fit<-lm(Expense~gender+house+income+Channel+age,data=mdat)
summary(fit)
```

## Correlation Plot

Check correlation plot for some of the variables:

```{r, message=FALSE}
library(caret)
library(corrplot)
# subset data
sdat<-subset(sim.dat,select=c("age","income","store_exp","online_exp","store_trans","online_trans" ))
# impute missiong values
imp<-preProcess(sdat,method="bagImpute")
sdat<-predict(imp,sdat)
# get correlation matrix
correlation<-cor(sdat)
# correlation plot
par(oma=c(2,2,2,2))
corrplot.mixed(correlation,order="hclust",tl.pos="lt",upper="ellipse")
```


