---
title: "Chapter 7 Model Tuning Strategy"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

In this notebook, we illustrate basic model tuning strategies.

# Load R Packages and Data

```{r, message = FALSE, warning=FALSE, results='hide'}
# install packages from CRAN
p_needed <- c('ggplot2','tidyr', 'caret', 'dplyr', 'lattice', 'proxy', 'caret')
packages <- rownames(installed.packages())
p_to_install <- p_needed[!(p_needed %in% packages)]
if (length(p_to_install) > 0) {
    install.packages(p_to_install)
}

lapply(p_needed, require, character.only = TRUE)
```

# Variance-Bias Trade-Off

In this section, we will show the variance-bias trade-off through a simulated example. We first simulate a dataset and plot the data.

```{r}
# randomly simulate some non-linear samples
x = seq(1, 10, 0.01) * pi
e = rnorm(length(x), mean = 0, sd = 0.2)
fx <- sin(x) + e + sqrt(x)
dat = data.frame(x, fx)
```

If we fit a linear regression model to this dataset, it will be a high bias model.

```{r}
ggplot(dat, aes(x, fx)) +
geom_point() +
geom_smooth(method = "lm", se = FALSE)
```

If we use a smoothing method to fit the data, it will be a high variance model.

```{r}
ggplot(dat, aes(x, fx)) + geom_smooth(span = 0.03)
```

To better illustrate the variance-bias trade-off, we now simulate four subsets of the original dataset and try to fit a smoothing model and a linear model. As the plots show, the smoothing model is more flexible to the training data, while the linear model is more robust to different subsets.

```{r}
# set random seed
set.seed(2016)
# sample part of the data to fit model sample 1
idx1 = sample(1:length(x), 100)
dat1 = data.frame(x1 = x[idx1], fx1 = fx[idx1])
p1 = ggplot(dat1, aes(x1, fx1)) + 
  geom_smooth(span = 0.03) + 
  geom_point()

# sample 2
idx2 = sample(1:length(x), 100)
dat2 = data.frame(x2 = x[idx2], fx2 = fx[idx2])
p2 = ggplot(dat2, aes(x2, fx2)) + 
  geom_smooth(span = 0.03) + 
  geom_point()

# sample 3
idx3 = sample(1:length(x), 100)
dat3 = data.frame(x3 = x[idx3], fx3 = fx[idx3])
p3 = ggplot(dat3, aes(x3, fx3)) + 
  geom_smooth(span = 0.03) + 
  geom_point()

# sample 4
idx4 = sample(1:length(x), 100)
dat4 = data.frame(x4 = x[idx4], fx4 = fx[idx4])
p4 = ggplot(dat4, aes(x4, fx4)) + 
  geom_smooth(span = 0.03) + 
  geom_point()

# Load multiplot function
source('http://bit.ly/2KeEIg9')

multiplot(p1, p2, p3, p4, cols = 2)
```


```{r}
p1 = ggplot(dat1, aes(x1, fx1)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  geom_point()

p2 = ggplot(dat2, aes(x2, fx2)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  geom_point()

p3 = ggplot(dat3, aes(x3, fx3)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  geom_point()

p4 = ggplot(dat4, aes(x4, fx4)) + 
  geom_smooth(method = "lm", se = FALSE) + 
  geom_point()

multiplot(p1, p2, p3, p4, cols = 2)
```

# Data Splitting and Resampling

Data splitting and resampling are fundamental techniques to build sound models for prediction.

## Data Splitting

The `createDataPartition()` function in `caret` package will return a balanced splitting based on assigned variable.


```{r}
# load data
sim.dat <- read.csv("http://bit.ly/2P5gTw4")
# set random seed to make sure reproducibility
set.seed(3456)
trainIndex <- createDataPartition(sim.dat$segment,
                                  p = 0.8,
                                  list = FALSE,
                                  times = 1)
head(trainIndex)
```


```{r}
# get training set
datTrain <- sim.dat[trainIndex, ]
# get test set
datTest <- sim.dat[-trainIndex, ]
```

Let's check the distribution of the two sets:

```{r, message=FALSE}
datTrain %>%
  dplyr::group_by(segment) %>%
  dplyr::summarise(count = n(),
            percentage = round(length(segment)/nrow(datTrain), 2))
```

```{r, message=FALSE}
datTest %>%
  dplyr::group_by(segment) %>%
  dplyr::summarise(count = n(),
            percentage = round(length(segment)/nrow(datTrain), 2))
```

An alternative way is to split data based on the predictors. The goal is to get a diverse subset from a dataset so that the sample is representative.

```{r}
# select variables
testing <- subset(sim.dat, select = c("age", "income"))
```

Random select 5 samples as initial subset (`start`) , the rest will be in `samplePool`: 

```{r}
set.seed(5)
# select 5 random samples
startSet <- sample(1:dim(testing)[1], 5)
start <- testing[startSet, ]
# save the rest in data frame 'samplePool'
samplePool <- testing[-startSet, ]
```

Use `maxDissim()` to select another 5 samples from `samplePool` that are as different as possible with the initical set `start`:

```{r}
selectId <- maxDissim(start, samplePool, obj = minDiss, n = 5)
minDissSet <- samplePool[selectId, ]
```

```{r}
selectId <- sample(1:dim(samplePool)[1], 5)
RandomSet <- samplePool[selectId, ]
```


```{r}
start$group <- rep("Initial Set", nrow(start))
minDissSet$group <- rep("Maximum Dissimilarity Sampling",
                        nrow(minDissSet))

RandomSet$group <- rep("Random Sampling",
                        nrow(RandomSet))

xyplot(age ~ income,
       data = rbind(start, minDissSet, RandomSet),
       grid = TRUE,
       group = group,
       auto.key = TRUE)
```

For time series data, random sampling is usually not the best way.

```{r}
# simulte AR(1) time series samples
timedata = arima.sim(list(order=c(1,0,0), ar=-.9), n=100)
# plot time series
plot(timedata, main=(expression(AR(1)~~~phi==-.9)))
```

```{r}
timeSlices <- createTimeSlices(1:length(timedata),
                               initialWindow = 36,
                               horizon = 12,
                               fixedWindow = T)
str(timeSlices,max.level = 1)
```

```{r}
# get result for the 1st training set
trainSlices <- timeSlices[[1]]
# get result for the 1st test set
testSlices <- timeSlices[[2]]
# check the index for the 1st training and test set
trainSlices[[1]]
```


```{r}
testSlices[[1]]
```

## Resampling

k-fold cross-validation is to partition the original sample into ð‘˜ equal size subsamples (folds).

```{r}
class <- sim.dat$segment
# creat k-folds
set.seed(1)
cv <- createFolds(class, k = 10, returnTrain = T)
str(cv)
```

Repeated training/test splits method is repeating the training/test set division on the original data.


```{r}
trainIndex <- createDataPartition(sim.dat$segment,
                                  p = .8,
                                  list = FALSE,
                                  times = 5)
dplyr::glimpse(trainIndex)
```
